---
title: "practice"
author: "youngtaek"
date: "2023-10-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(here)
source(here('lib','get_lib.R'))
```

# Practice data analysis

## Wk 1 practice

### How to create amtrix and vector in R

$$\begin{aligned}
\mathbb{A}\cdot\vec{x} = \vec{b}
\end{aligned}$$

$$\begin{aligned}
\vec{v}
\end{aligned}$$

```{r}
a1 <- matrix(c(3, 0, -1, -5, 2, 4), nrow = 1, byrow = T)
print(a1)
```

```{r}
a2 <- matrix(c(3, 0, -1, 5, 2, 4), nrow = 1, byrow = T)

A <- rbind(a1, a2)
print(A)
```

```{r}
x <- c(1, 2, 3, 4, 5, 6)
x
b <- x/Norm(x)
Norm(b)
A%*%x
```

$$\begin{aligned}
\mathbb{B}\cdot\vec{x}_B + \mathbb{N}\cdot\vec{x}_N = \mathbb{A}\cdot\vec{x}
\end{aligned}$$

### Creating saple vector

```{r}
#randomly selets number
a <- sample(-5:5, replace = TRUE, 12)
a
length(a)
A <- matrix(a, ncol = 4, byrow = TRUE)
A
```

###Plotting
####squirrels
```{r}
squirrel_subset <- fread(here('data/squirrels_subset.csv'))

ggplot(data = squirrel_subset) + aes(x = long, y = lat) + geom_point()
```

#### Color by age

```{r}
ggplot(data = squirrel_subset) + aes(x = long, y = lat, color = age) + geom_point()
```

#### Make Every Point blue
```{r}
ggplot(data = squirrel_subset) + aes(x = long, y = lat) + geom_point(color = 'blue')
```

#### Color by the Fur Color
```{r}
ggplot(data = squirrel_subset) + aes(x = long, y = lat, color = primary_fur_color) + geom_point()
```

#### Put onto non-euclidian space
```{r}
ggplot(data = squirrel_subset) + aes(x = long, y = lat, color = primary_fur_color) + coord_quickmap() + geom_point()
```

```{r}
squirrel_subset_by_color <- squirrel_subset %>%  group_by(primary_fur_color) %>%  summarise(count_by_color = n())

plot_col <- squirrel_subset_by_color %>%  ggplot() + aes(x = primary_fur_color, y = count_by_color) + geom_col()

plot_bar <- squirrel_subset %>%  ggplot() + aes(x = primary_fur_color) + geom_bar()

squirrel_subset_by_color <- squirrel_subset %>% mutate(date_f = as.Date.character(date, format = '%m%d%Y')) %>% group_by(date_f, primary_fur_color) %>%  summarise(count_of_squirrels = n())
```

```{r}
squirrel_subset_by_color %>%  ggplot() + aes(x = date_f, y = count_of_squirrels, color = primary_fur_color) + stat_smooth(se = FALSE) + labs(title = 'Decreasing Count of Squirrels Through Time', subtitle = 'Moving average smoother estimate', x = 'Date of Observation', y = 'Count of Squirrels', color = 'Primary Fur Color') + coord_cartesian(xlim = c(as.Date.character('2018-10-08'), as.Date.character('2018-10-15')))
```


## Wk 2 practice

### Creating interactive plot
```{r}
hprice1 <- wooldridge::hprice1
glimpse(hprice1)

p1 <- hprice1 %>% ggplot(aes(x = colonial, y = price)) + geom_point()
p1

ggplotly(p1)
```

### filter 관련
```{r}
agencies <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-15/agencies.csv')
glimpse(agencies)
```

```{r}
launches <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-15/launches.csv')

glimpse(launches)
```

```{r}
filter(launches, launch_year < 1968, launch_year > 1966, agency == 'US')
```

```{r}
launches %>%
  filter(launch_year < 1968) %>% 
  filter(launch_year > 1966) %>% 
  filter(agency == 'US' | agency == 'SU')
```

### arrange()
```{r}
agencies %>% 
  arrange(desc(state_code))
```

### select()
```{r}
launches %>% 
  select(contains('agency'), contains('launch')) %>% 
  view()
```

### summarise()
```{r}
agencies %>% 
  summarize(
    average_launches = mean(count),
    var_launches = var(count),
    number_of_agencies = length(unique(agencies)),
    count_log = log(count)
  ) %>%  head()
```

### mutate()
```{r}
agencies %>% 
  mutate(count_log = log(count)) %>%  head() %>% 
  select(contains('count'))
```

## Wk 3 practice
```{r}
d <- data.frame(
  id = 1:15,
  individual = rep(LETTERS[1:5], each = 3),
  group = rep(c('Group One', 'Group Two'), times = c(9, 6)), time = rep(1:3, times = 5), value = (1:15)^2
)
d

d <- d %>%
  arrange(sample(id))
d %>%  group_by(group, time) %>% 
  summarise(
    value_average = mean(value)
  )
```

### how to read data from github
```{r}
agencies <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-15/agencies.csv')
launches <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-15/launches.csv')

launches %>% 
  group_by(state_code, launch_year) %>% 
  summarise(total_launches = n()) %>% 
  arrange(desc(total_launches))
```

### Graphing
#### patchwork and here
```{r}
squirrel_subset <- fread(here("data", "squirrels_subset.csv"))
```

```{r}
squirrel_scatter <- squirrel_subset %>% 
  ggplot() +
  aes(x = long, y = lat) +
  geom_point()
squirrel_scatter
head(squirrel_subset)

squirrel_long_histogram <- squirrel_subset %>% 
  ggplot() +
  aes(x = long) +
  geom_histogram()
squirrel_long_histogram
squirrel_long_density <- squirrel_subset %>%
  ggplot() +
  aes(x = long) +
  geom_density()
squirrel_long_density

squirrel_long_histogram / squirrel_long_density
```

#### change the format of the data
```{r}
squirrel_subset <- squirrel_subset %>% 
  mutate(
    date_f = as.Date.character(date, format = '%m%d%Y')
)
```

```{r}
squirrel_subset_by_color <- squirrel_subset %>% 
  group_by(primary_fur_color) %>% 
  summarise(count_by_color = n())
```

```{r}
squirrel_subset %>% 
  ggplot() +
  aes(x = date_f) +
  geom_histogram() +
  facet_wrap(vars(primary_fur_color), nrow = 3)
```

```{r}
plot_col <- squirrel_subset_by_color %>% 
  ggplot() +
  aes(x = primary_fur_color, y = count_by_color) + 
  geom_col()

plot_bar <- squirrel_subset %>%
  ggplot() +
  aes(x = primary_fur_color) +
  geom_bar()

plot_col | plot_bar
```

## Wk 4 practice
### Invertible Linear Transformation

$$\begin{aligned}
\mathbb{A}\vec{x} = \vec{b}
\end{aligned}$$

$$\begin{aligned}
\mathbb{A}^{-1}\vec{b} = \vec{x}
\end{aligned}$$

```{r}
r1 <- c(0, 1, -4, 5)
r2 <- c(2, -3, 2, 7)
r3 <- c(5, -8, 9, 10)
r4 <- c(15, -18, 19, 10)
A <- rbind(r1, r2, r3, r4)
print(A)

Rank(A)

rref(A)

x <- c(3, 6, 9)
x
T <- A
T

b <- T%*%x
```

